Describe Participants: Characteristics, Diet, and Cardiometabolic Risk

### 2021-12-01 Rmarkdown/analysis/Characteristics_Diet_CardiometabolicRisk_211111.Rmd
Add dietary choline to analysis

# Make TMAO log variable
phen$tmao_log <- log(phen$tmao)
```

@@

# Combine food with phenotype data
full <- merge(phen, food, by = 0)
row.names(full) <- full[,"Row.names"]
full <- subset(full, select = -c(Row.names))
```

@@

```{r Foods to keep, results='hide'}
transformed_keep_tbl <- c("sex", 
#"sex_numeric",
"kcal_tnfs", 
"age", 
"cystatinc_bd1", 
"ethnicity",
"tmao", 
"tmao_log", 
#"TMAO.0",
#"TMAO.30",
#"TMAO.180",
#"TMAO.360",
#"lnTMAO.0",
#"lnTMAO.30",
#"lnTMAO.180",
#"lnTMAO.360",
#"TMAO_0_30",
#"TMAO_0_180",
#"TMAO_0_360",
#"Ind.SO4",
#"p.Cresol.SO4",
#"TMAO_quart_num",
#"T.pf_mps_totals_sqrRt",
"T.pf_meat_totals_sqrRt",
"T.pf_curedmeat_totals_sqrRt",
"T.pf_poult_totals_sqrRt",
"T.pf_seafd_hi_totals_Binary",
"T.pf_seafd_low_totals_Binary",
"T.pf_eggs_totals_sqrRt",
"T.pf_nutsds_totals_sqrRt",
"T.pf_legumes_totals_sqrRt",
"T.pf_total_totals_natLog",
#"T.pf_soy_totals_Binary",
"T.v_drkgr_totals_sqrRt",
"T.v_redor_total_totals_sqrRt",
#"T.v_redor_tomato_totals_sqrRt",
#"T.v_redor_other_totals_sqrRt",
"T.v_starchy_total_totals_sqrRt",
#"T.v_starchy_potato_totals_sqrRt",
#"T.v_starchy_other_totals_sqrRt",
"T.v_other_totals_sqrRt",
"T.v_total_totals_sqrRt",
#"T.v_legumes_total_sqrRt",
"T.g_whole_totals_sqrRt",
"T.g_refined_totals_sqrRt",
"g_total_totals",
"T.d_total_totals_sqrRt",
"T.d_milk_totals_sqrRt",
"T.d_cheese_totals_sqrRt",
"T.d_yogurt_totals_Binary"
)
```

@@

Make TMAO food intake by quantile table
```{r Food Intake by TMAO quantile}
# TMAO table by age bin - females
factorVars <- c("age_cat", "tmao_quantile")
vars <- c("pf_meat_totals", 
          "pf_curedmeat_totals",
          "pf_poult_totals",
          "pf_seafd_hi_totals",
          "pf_seafd_low_totals",
          "pf_eggs_totals",
          "pf_nutsds_totals",
          "pf_legumes_totals",
          "pf_total_totals",
          #"pf_soy_totals_Binary",
          "v_drkgr_totals",
          "v_redor_total_totals",
          #"v_redor_tomato_totals",
          #"v_redor_other_totals",
          "v_starchy_total_totals",
          #"v_starchy_potato_totals",
          #"v_starchy_other_totals",
          "v_other_totals",
          "v_total_totals",
          #"v_legumes_total",
          "g_whole_totals",
          "g_refined_totals",
          "g_total_totals",
          "d_total_totals",
          "d_milk_totals",
          "d_cheese_totals",
          "d_yogurt_totals",
          "hei_asa24_totalscore",
          "hei_asa24_totalveg",
          "hei_asa24_sfat",
          "hei_asa24_",
          "age",
          "ethnicity" ,
          "bmi_final",
          "tmao",
          "choline",
          "betaine",
          "carnitine")
table_food.fq <- CreateTableOne(vars = vars,
                                    strata = "tmao_quantile",
                                    data = full[full$sex == "Female",],
                                    factorVars = factorVars)
table_food.fq

@@

TMAO tertile 
Make TMAO food intake by tertile table
```{r Food Intake by TMAO tertile}
# TMAO table by age bin - females
factorVars <- c("age_cat", "tmao_quantile")
vars <- c("pf_meat_totals", 
          "pf_curedmeat_totals",
          "pf_poult_totals",
          "pf_seafd_hi_totals",
          "pf_seafd_low_totals",
          "pf_eggs_totals",
          "pf_nutsds_totals",
          "pf_legumes_totals",
          "pf_total_totals",
          #"pf_soy_totals_Binary",
          "v_drkgr_totals",
          "v_redor_total_totals",
          #"v_redor_tomato_totals",
          #"v_redor_other_totals",
          "v_starchy_total_totals",
          #"v_starchy_potato_totals",
          #"v_starchy_other_totals",
          "v_other_totals",
          "v_total_totals",
          #"v_legumes_total",
          "g_whole_totals",
          "g_refined_totals",
          "g_total_totals",
          "d_total_totals",
          "d_milk_totals",
          "d_cheese_totals",
          "d_yogurt_totals",
          "hei_asa24_totalscore",
          "hei_asa24_totalveg",
          "hei_asa24_sfat",
          "hei_asa24_",
          "age",
          "ethnicity" ,
          "bmi_final",
          "tmao",
          "choline",
          "betaine",
          "carnitine")
table_food.ft <- CreateTableOne(vars = vars,
                                    strata = "tmao_tertile",
                                    data = full[full$sex == "Female",],
                                    factorVars = factorVars)
table_food.ft

@@

```{r}
full$low_mps_meat <- ifelse(full$pf_mps_total_totals <= median(full$pf_mps_total_totals), "low", "high")
full$low_mps_meat <- ifelse(full$pf_mps_total_totals <= 9.091, "low", "high")
summary(lm(tmao_log ~ low_mps_meat + sex*age + cystatinc_bd1 + kcal_tnfs, full))
```

>>>>

# Make TMAO log variable
phen$tmao_log <- log(phen$tmao)

```

@@

# Combine food with phenotype data
full <- merge(phen, food, by = 0)
row.names(full) <- full[,"Row.names"]
full <- subset(full, select = -c(Row.names))

# Make ASA24 choline log variable
full$choln_tnfs_log <- log(full$choln_tnfs)
```

@@

```{r Foods to keep, results='hide'}
transformed_keep_tbl <- c("sex", 
#"sex_numeric",
"kcal_tnfs", 
"age", 
"cystatinc_bd1", 
"ethnicity",
"tmao", 
"tmao_log", 
#"TMAO.0",
#"TMAO.30",
#"TMAO.180",
#"TMAO.360",
#"lnTMAO.0",
#"lnTMAO.30",
#"lnTMAO.180",
#"lnTMAO.360",
#"TMAO_0_30",
#"TMAO_0_180",
#"TMAO_0_360",
#"Ind.SO4",
#"p.Cresol.SO4",
#"TMAO_quart_num",
#"T.pf_mps_totals_sqrRt",
"T.pf_meat_totals_sqrRt",
"T.pf_curedmeat_totals_sqrRt",
"T.pf_poult_totals_sqrRt",
"T.pf_seafd_hi_totals_Binary",
"T.pf_seafd_low_totals_Binary",
"T.pf_eggs_totals_sqrRt",
"T.pf_nutsds_totals_sqrRt",
"T.pf_legumes_totals_sqrRt",
"T.pf_total_totals_natLog",
#"T.pf_soy_totals_Binary",
"T.v_drkgr_totals_sqrRt",
"T.v_redor_total_totals_sqrRt",
#"T.v_redor_tomato_totals_sqrRt",
#"T.v_redor_other_totals_sqrRt",
"T.v_starchy_total_totals_sqrRt",
#"T.v_starchy_potato_totals_sqrRt",
#"T.v_starchy_other_totals_sqrRt",
"T.v_other_totals_sqrRt",
"T.v_total_totals_sqrRt",
#"T.v_legumes_total_sqrRt",
"T.g_whole_totals_sqrRt",
"T.g_refined_totals_sqrRt",
"g_total_totals",
"T.d_total_totals_sqrRt",
"T.d_milk_totals_sqrRt",
"T.d_cheese_totals_sqrRt",
"T.d_yogurt_totals_Binary",
"choln_tnfs_log"
)
```

@@

Make TMAO food intake by quantile table
```{r Food Intake by TMAO quantile}
# TMAO table by age bin - females
factorVars <- c("age_cat", "tmao_quantile")
vars <- c("pf_meat_totals", 
          "pf_curedmeat_totals",
          "pf_poult_totals",
          "pf_seafd_hi_totals",
          "pf_seafd_low_totals",
          "pf_eggs_totals",
          "pf_nutsds_totals",
          "pf_legumes_totals",
          "pf_total_totals",
          #"pf_soy_totals_Binary",
          "v_drkgr_totals",
          "v_redor_total_totals",
          #"v_redor_tomato_totals",
          #"v_redor_other_totals",
          "v_starchy_total_totals",
          #"v_starchy_potato_totals",
          #"v_starchy_other_totals",
          "v_other_totals",
          "v_total_totals",
          #"v_legumes_total",
          "g_whole_totals",
          "g_refined_totals",
          "g_total_totals",
          "d_total_totals",
          "d_milk_totals",
          "d_cheese_totals",
          "d_yogurt_totals",
          "hei_asa24_totalscore",
          "hei_asa24_totalveg",
          "hei_asa24_sfat",
          "hei_asa24_",
          "choln_tnfs",
          "age",
          "ethnicity" ,
          "bmi_final",
          "tmao",
          "choline",
          "betaine",
          "carnitine")
table_food.fq <- CreateTableOne(vars = vars,
                                    strata = "tmao_quantile",
                                    data = full[full$sex == "Female",],
                                    factorVars = factorVars)
table_food.fq

@@

TMAO tertile 
Make TMAO food intake by tertile table
```{r Food Intake by TMAO tertile}
# TMAO table by age bin - females
factorVars <- c("age_cat", "tmao_quantile")
vars <- c("pf_meat_totals", 
          "pf_curedmeat_totals",
          "pf_poult_totals",
          "pf_seafd_hi_totals",
          "pf_seafd_low_totals",
          "pf_eggs_totals",
          "pf_nutsds_totals",
          "pf_legumes_totals",
          "pf_total_totals",
          #"pf_soy_totals_Binary",
          "v_drkgr_totals",
          "v_redor_total_totals",
          #"v_redor_tomato_totals",
          #"v_redor_other_totals",
          "v_starchy_total_totals",
          #"v_starchy_potato_totals",
          #"v_starchy_other_totals",
          "v_other_totals",
          "v_total_totals",
          #"v_legumes_total",
          "g_whole_totals",
          "g_refined_totals",
          "g_total_totals",
          "d_total_totals",
          "d_milk_totals",
          "d_cheese_totals",
          "d_yogurt_totals",
          "hei_asa24_totalscore",
          "hei_asa24_totalveg",
          "hei_asa24_sfat",
          "hei_asa24_",
          "choln_tnfs",
          "age",
          "ethnicity" ,
          "bmi_final",
          "tmao",
          "choline",
          "betaine",
          "carnitine")
table_food.ft <- CreateTableOne(vars = vars,
                                    strata = "tmao_tertile",
                                    data = full[full$sex == "Female",],
                                    factorVars = factorVars)
table_food.ft

@@

"waistavg_cm", "sysbpavgv2", "diabpavgv2", "endo_rhi", "endo_al75", "hdl_bd1", "ldl_bd1", "chol_bd1", "tg_bd1_ln", "glc_bd1_ln","tnfa_bd1_ln", "il6_bd1_ln", "crp_bd1_ln",

BS section below...
```{r}
full$low_mps_meat <- ifelse(full$pf_mps_total_totals <= median(full$pf_mps_total_totals), "low", "high")
full$low_mps_meat <- ifelse(full$pf_mps_total_totals <= 9.091, "low", "high")
summary(lm(glc_bd1_ln ~  tmao_log + sex*age + cystatinc_bd1, full[full$low_mps_meat == "high",]))
summary(lm(glc_bd1_ln ~  tmao_log + sex*age + cystatinc_bd1, full[full$low_mps_meat == "high",]))
```
